Reasonably Programmable Literal Notation

### 2017-10-28 pldi18/syntax-pldi18.tex
figure tweaks

%% Listings
\definecolor{mygray}{rgb}{0.75,0.75,0.75}
\usepackage{listings}
\lstset{tabsize=2, 
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont, 
commentstyle=\itshape\ttfamily\color{gray},
stringstyle=\ttfamily\color{red},
mathescape=false,escapechar=~,
numbers=left, numberstyle=\scriptsize\color{mygray}, language=ML,moredelim=[il][\sffamily]{?},showspaces=false,showstringspaces=false,xleftmargin=0pt, numbersep=-6pt, morekeywords=[1]{spliced,tyfam,opfam,let,fn,val,def,casetype,objtype,metadata,of,*,string,datatype,new,toast,syntax,module,switch,where,using,import,for,ana,syn,opcon,tycon,metasignature,metamodule,metasig,metamod,static,at,by,tycase,mod,macro,match,pattern,in,patterns,expressions,implicit,forall,exptsm,pattsm},deletekeywords={double},classoffset=0, xleftmargin=-10pt, 
aboveskip=4pt,belowskip=2pt,
moredelim=**[is][\color{red}]{SSTR}{ESTR},
moredelim=**[is][\color{OliveGreen}]{SHTML}{EHTML},
moredelim=**[is][\color{purple}]{SCSS}{ECSS},
moredelim=**[is][\color{brown}]{SSQL}{ESQL},
moredelim=**[is][\color{orange}]{SCOLOR}{ECOLOR},
moredelim=**[is][\color{magenta}]{SPCT}{EPCT}, 
moredelim=**[is][\color{gray}]{SNAT}{ENAT}, 
moredelim=**[is][\color{OliveGreen}]{SURL}{EURL},
moredelim=**[is][\color{SeaGreen}]{SQT}{EQT},
moredelim=**[is][\color{Periwinkle}]{SGRM}{EGRM},
moredelim=**[is][\color{YellowGreen}]{SID}{EID},
moredelim=**[is][\color{Sepia}]{SUS}{EUS},
deletestring=[d]{"},
}
\lstset{morecomment=[n]{/*}{*/}}

@@

%\footnote{In general, spliced expressions might themselves apply TLMs, in which case the convention is to use a distinct color for unspliced segments at each depth. For example, if strings were not primitive, we might write \li{`SURL<body>\{EURLheading ($str "SSTRWorld!ESTR")SURL\} ...</body>EURL`}.}

@@

\vspace{-4px}
\newcommand{\seTLMsSec}{Simple Expression TLMs (seTLMs)}
\section{\protect\seTLMsSec}
\label{sec:setsms}
\vspace{-2px}
We will start in this section by informally describing TLMs in Reason. It will suffice in this section to consider only \emph{simple expression TLMs (seTLMs)}, i.e. TLMs for abbreviating expressions of a single specified type. In particular, we will consider an seTLM named \li{$html} that implements Ur/Web style HTML literal notation for expressions of the \li{html} type that is defined in Figure \ref{fig:html-type-def}.

\begin{figure}[ht!]
\vspace{-10px}
\begin{lstlisting}[numbers=none]
  type html = BodyElement(list(html_attr), list(html))
            | H1Element(list(html_attr), list(html))
            | TextNode(string) | /* ... */;
\end{lstlisting}
\vspace{-10px}
\caption{The \li{html} type, which classifies encodings of HTML data. The type \li{html_attr} is also straightforward and the type function \li{list} is standard, so neither are shown. Strings are primitive.}
\label{fig:html-type-def}
\vspace{-10px}
\end{figure}

@@

\begin{figure}[ht!]
\vspace{-12px}
\begin{lstlisting}[mathescape=|]
  syntax $html at html by static {
    fun(b : body) : parse_result(proto_expr) => 
      /* ... Ur/Web-style HTML syntax parser ... */
  };
\end{lstlisting}
\vspace{-15px}
\caption{The definition of the seTLM named \li{$html}.}
\label{fig:html-tlm-def}
\vspace{-10px}
\end{figure}

@@

\begin{figure}[t!]
\vspace{-2px}
\begin{lstlisting}
  fun heading first_name = $html [|SURL<h1>Hello, {[EURLfirst_nameSURL]}!</h1>EURL|]
  val body = $html [|SURL<body>{EURLheading "World!"SURL} ...</body>EURL|]
\end{lstlisting}
\vspace{-10px}
\caption{Two examples of the \li{$html} TLM being applied. %In each case, literal body, between backticks, is initially left unparsed according to the language's context-free syntax. %The applied TLM determines a segmentation and expansion during the typed expansion phase, which generalizes the usual typing phase.
}
\vspace{-12px}
\label{fig:first-tsm-example}
\end{figure}

@@

\begin{figure}
\vspace{-2px}
\begin{lstlisting}[numbers=none]
type body = string;
type segment = {startIdx : int, endIdx : int}; /* inclusive */
type parse_result('a) = ParseError { msg : string, loc : segment }
                      | Success('a);
type var_t = string;
type proto_typ = TyVar(var_t)
               | Arrow(proto_typ, proto_typ)
               | StringTy
               | /* ... */ 
               | SplicedT(segment);
type proto_expr = Var(var_t)
                | Fn(var_t, proto_typ, proto_expr)
                | Ap(proto_expr, proto_expr)
                | /* ... */ 
                | SplicedE(segment, proto_typ);
\end{lstlisting}
\vspace{-10px}
\caption[Definitions of various types used by TLM definitions.]{Definitions of various types used by TLM definitions. We assume that these are available ambiently.}
\label{fig:indexrange-and-parseresult}
\label{fig:candidate-exp-Reason}
\vspace{-12px}
\end{figure}

>>>>

%% Listings
\definecolor{mygray}{rgb}{0.75,0.75,0.75}
\usepackage{listings}
\lstset{tabsize=2, 
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont, 
commentstyle=\itshape\ttfamily\color{gray},
stringstyle=\ttfamily\color{red},
mathescape=false,escapechar=~,
numbers=left, numberstyle=\scriptsize\color{mygray}, language=ML,moredelim=[il][\sffamily]{?},showspaces=false,showstringspaces=false,xleftmargin=0pt, numbersep=-6pt, morekeywords=[1]{try,spliced,tyfam,opfam,let,fn,val,def,casetype,objtype,metadata,of,*,string,datatype,new,toast,syntax,module,switch,where,using,import,for,ana,syn,opcon,tycon,metasignature,metamodule,metasig,metamod,static,at,by,tycase,mod,macro,match,pattern,in,patterns,expressions,implicit,forall,exptsm,pattsm},deletekeywords={double},classoffset=0, xleftmargin=-10pt, 
aboveskip=4pt,belowskip=2pt,
moredelim=**[is][\color{red}]{SSTR}{ESTR},
moredelim=**[is][\color{OliveGreen}]{SHTML}{EHTML},
moredelim=**[is][\color{purple}]{SCSS}{ECSS},
moredelim=**[is][\color{brown}]{SSQL}{ESQL},
moredelim=**[is][\color{orange}]{SCOLOR}{ECOLOR},
moredelim=**[is][\color{magenta}]{SPCT}{EPCT}, 
moredelim=**[is][\color{gray}]{SNAT}{ENAT}, 
moredelim=**[is][\color{OliveGreen}]{SURL}{EURL},
moredelim=**[is][\color{SeaGreen}]{SQT}{EQT},
moredelim=**[is][\color{Periwinkle}]{SGRM}{EGRM},
moredelim=**[is][\color{YellowGreen}]{SID}{EID},
moredelim=**[is][\color{Sepia}]{SUS}{EUS},
deletestring=[d]{"},
}
\lstset{morecomment=[n]{/*}{*/}}

@@

%\footnote{In general, spliced expressions might themselves apply TLMs, in which case the convention is to use a distinct color for unspliced segments at each depth. For example, if strings were not primitive, we might write \li{`SURL<body>\{EURLheading ($str "SSTRWorld!ESTR")SURL\} ...</body>EURL`}.}
\begin{figure*}[t]

\begin{subfigure}[t]{0.48\textwidth}
\begin{lstlisting}
  type html_attrs = list(string, string)
  type html = BodyElement(html_attrs, list(html))
            | H1Element(html_attrs, list(html))
            | TextNode(string) 
            | /* ... */;
\end{lstlisting}
\vspace{-5px}
\caption{The \li{html} type, which classifies encodings of HTML data.}
\label{fig:html-type-def}
\end{subfigure}
\hfill
\begin{subfigure}[t]{0.48\textwidth}
\begin{lstlisting}[mathescape=|]
  syntax $html at html by static {
    fun(b : body) : parse_result(proto_expr) => 
      try (parse_html b |> html_to_ast |> Success) {
      | InvalidHTML msg loc => ParseError msg loc }
  };
\end{lstlisting}
\vspace{-5px}
\caption{The definition of the \li{$html} TLM. It calls into an existing HTML parser, then converts the result to an AST, extracting the spliced segments as it goes. Figure \ref{fig:candidate-exp-Reason} defines the  TLM-related types.}
\label{fig:html-tlm-def}
\end{subfigure}

\begin{subfigure}[t]{\textwidth}
\begin{lstlisting}
  let heading first_name = $html [|SURL<h1>Hello, {[EURLfirst_nameSURL]}!</h1>EURL|]
  let body = $html [|SURL<body>{EURLheading "World!"SURL} ...</body>EURL|]
\end{lstlisting}
\vspace{-5px}
\caption{Two examples of the \li{$html} TLM being applied. Compare to Ur/Web's built in HTML literal notation from Figure \ref{fig:urweb}.  %In each case, literal body, between backticks, is initially left unparsed according to the language's context-free syntax. %The applied TLM determines a segmentation and expansion during the typed expansion phase, which generalizes the usual typing phase.
}
\label{fig:first-tsm-example}
\end{subfigure}
\vspace{-8px}
\caption{Case Study: HTML literals}
\vspace{-5px}
\end{figure*}

@@

\vspace{-4px}
\newcommand{\seTLMsSec}{Simple Expression TLMs (seTLMs)}
\section{\protect\seTLMsSec}
\label{sec:setsms}
\vspace{-2px}

We will start in this section by informally describing TLMs in Reason. It will suffice in this section to consider only \emph{simple expression TLMs (seTLMs)}, i.e. TLMs for abbreviating expressions of a single specified type. In particular, we will consider an seTLM named \li{$html} that implements Ur/Web style HTML literal notation for expressions of the \li{html} type that is defined in Figure \ref{fig:html-type-def}.

@@

@@

@@

\begin{figure}
\vspace{-2px}
\begin{lstlisting}[numbers=none,xleftmargin=0pt]
type body = string;
type segment = {startIdx : int, endIdx : int};
type parse_result('a) 
  = ParseError { msg : string, loc : segment }
  | Success('a);
type var_t = string;
type proto_typ = TyVar(var_t)
               | Arrow(proto_typ, proto_typ)
               | StringTy
               | /* ... */ 
               | SplicedT(segment);
type proto_expr = Var(var_t)
                | Fn(var_t, proto_typ, proto_expr)
                | Ap(proto_expr, proto_expr)
                | /* ... */ 
                | SplicedE(segment, proto_typ);
\end{lstlisting}
\vspace{-10px}
\caption[Definitions of various types used by TLM definitions.]{Definitions of various types available ambiently to TLM definitions.}
\label{fig:indexrange-and-parseresult}
\label{fig:candidate-exp-Reason}
\vspace{-12px}
\end{figure}
